% bibliography.tex
\begin{thebibliography}{99} % Adjust 99 if you have more than 99 entries

% Chapter A Refs
\bibitem{deberg2008}
M. de Berg, O. Cheong, M. van Kreveld, M. Overmars.
\textit{Computational Geometry: Algorithms and Applications (3rd ed.)}.
Springer-Verlag, 2008.
\label{ref:deberg2008}

\bibitem{cpalgorithms_basic_geom}
CP-Algorithms. \textit{Basic Geometry}.
Retrieved from \url{https://cp-algorithms.com/geometry/basic-geometry.html}
\label{ref:cpalgo_basic_geom}

% Chapter B Refs
\bibitem{goodman_orourke_polygons} % More specific for polygons
J. E. Goodman, J. O'Rourke (Eds.).
\textit{Handbook of Discrete and Computational Geometry (3rd ed.)} (Chapter on Polygons).
CRC Press, 2017.
\label{ref:goodman_orourke_polygons}

\bibitem{topcoder_geom_p2}
TopCoder Tutorial. \textit{Geometry Concepts Part 2 - Polygons, Area, Centroid, Pick's Theorem}.
(URL placeholder: Search TopCoder tutorials, e.g., \url{https://community.topcoder.com/tc?module=Static&d1=tutorials&d2=geometry2})
\label{ref:topcoder_geom_p2}

% Chapter C Refs
\bibitem{ericson_collision}
C. Ericson.
\textit{Real-Time Collision Detection} (Chapters on intersection tests, SAT).
Morgan Kaufmann, 2005.
\label{ref:ericson_collision}

\bibitem{nifty_sat}
Nifty Game Dev. \textit{Separating Axis Theorem (SAT) Explained}.
(URL placeholder: \url{https://www.niftygamedevelopment.com/separating-axis-theorem-sat-explained/} or similar visual explanation)
\label{ref:nifty_sat}

% Chapter D Refs
\bibitem{sedgewick_wayne_hulls} % More specific for hulls
R. Sedgewick, K. Wayne.
\textit{Algorithms (4th ed.)} (Section on Convex Hulls).
Addison-Wesley Professional, 2011.
\label{ref:sedgewick_wayne_hulls}

\bibitem{topcoder_srm_hull_tutorial} % General tutorial, SRM 2 not specified, but a common topic
TopCoder SRM Convex Hull Problem Tutorial (e.g., SRM 202).
(URL placeholder: Search TopCoder SRM editorials or tutorials on Convex Hulls)
\label{ref:topcoder_srm_hull_tutorial}

% Chapter E Refs
\bibitem{kactl_linecontainer}
KTH Algorithm Competition Template Library (KACTL). \textit{LineContainer.h}.
Retrieved from \url{https://github.com/kth-competitive-programming/kactl/blob/main/content/data-structures/LineContainer.h}
\label{ref:kactl_linecontainer}

\bibitem{pegwiki_cht}
G. Šinkevičius. \textit{Convex Hull Trick} (PEGWiki or CMU CS).
Retrieved from (e.g., \url{https://www.cs.cmu.edu/~ginsca/convex_hull_trick.txt} or common PEGWiki mirrors)
\label{ref:pegwiki_cht}

% Chapter F Refs
\bibitem{maspypy_slope_trick_blog}
maspypy. (Relevant blog post on Slope Trick).
(URL placeholder: Search maspypy's blog, e.g., \url{https://maspypy.github.io/})
\label{ref:maspypy_slope_trick_blog}

\bibitem{olphe_slope_trick_blog}
olphe. (Relevant blog post on Slope Trick).
(URL placeholder: Search olphe's blog, e.g., \url{https://olphe1.github.io/})
\label{ref:olphe_slope_trick_blog}

\bibitem{usaco_guide_slope_trick}
USACO Guide. \textit{Slope Trick}.
Retrieved from \url{https://usaco.guide/adv/slope}
\label{ref:usaco_guide_slope_trick}

% Chapter G Refs
\bibitem{fredman1979}
M. L. Fredman. (1979). "The Gaps Between Consecutive Primes and Related Problems." % Note: This is likely a misattribution for QI, Fredman's work on QI for DP is more like "On the decision tree complexity of the shortest path problems" or related to speeding up DP. The primary QI is from Yao.
% For QI in DP, often F. F. Yao's work is cited. "Efficient dynamic programming using quadrangle inequalities." (1980)
(More appropriate reference for QI/Knuth opt: Yao, F. F. (1980). Efficient dynamic programming using quadrangle inequalities. STOC '80.)
\label{ref:fredman1979_or_yao1980} % Placeholder, needs correct citation for QI

\bibitem{knuth1971_obst}
D. E. Knuth. (1971). "Optimum binary search trees." \textit{Acta Informatica}.
\label{ref:knuth1971_obst}

\bibitem{cpalgorithms_knuth_opt}
CP-Algorithms. \textit{Knuth's Optimization}.
Retrieved from \url{https://cp-algorithms.com/dynamic_programming/knuth-optimization.html}
\label{ref:cpalgorithms_knuth_opt}

% Chapter H Refs
\bibitem{deberg_ch2_ch10} % de Berg again, specific chapters
M. de Berg, et al. \textit{Computational Geometry: Algorithms and Applications (3rd ed.)} (Chapter 2: Line Segment Intersection, Chapter 10: More Geometric Data Structures).
(See \Cref{ref:deberg2008})
\label{ref:deberg_ch2_ch10}

\bibitem{stanford_cs166_sweep}
Stanford CS166 Notes. \textit{Sweep Algorithms}.
(URL placeholder: Search for Stanford CS166 lecture notes on sweep-line algorithms)
\label{ref:stanford_cs166_sweep}

% Chapter I Refs
\bibitem{samet_spatial_ds}
H. Samet.
\textit{Foundations of Multidimensional and Metric Data Structures}.
Morgan Kaufmann, 2006.
\label{ref:samet_spatial_ds}

\bibitem{kactl_spatial_ds}
KTH Algorithm Competition Template Library (KACTL). (e.g., KDTree.h).
(URL placeholder: \url{https://github.com/kth-competitive-programming/kactl/tree/main/content/data-structures})
\label{ref:kactl_spatial_ds}

% Chapter J Refs
\bibitem{megiddo_parametric_search}
N. Megiddo. (Various papers on Parametric Search, e.g., "Applying parallel computation algorithms in the design of serial algorithms" (1983)).
(URL placeholder: Search for Nimrod Megiddo's papers on parametric search)
\label{ref:megiddo_parametric_search}

\bibitem{topcoder_srm_parametric_search}
TopCoder SRM Editorials. (Problems solvable by parametric search).
(General reference, specific SRM editorials would be better if known)
\label{ref:topcoder_srm_parametric_search}

\bibitem{alien_trick_blogs}
Blog posts on "Alien Trick" (e.g., by scott\_wu, pashka).
(URL placeholder: Search for "Alien Trick competitive programming blog")
\label{ref:alien_trick_blogs}

% Chapter K Refs
\bibitem{deberg_arrangements_voronoi} % de Berg again
M. de Berg, et al. \textit{Computational Geometry: Algorithms and Applications (3rd ed.)} (Chapters on Arrangements, Duality, Voronoi, Delaunay).
(See \Cref{ref:deberg2008})
\label{ref:deberg_arrangements_voronoi}

\bibitem{orourke_comp_geom_c}
J. O'Rourke. (1998). \textit{Computational Geometry in C (2nd ed.)}.
Cambridge University Press.
\label{ref:orourke_comp_geom_c}

% Add any other necessary bibliography items here.

\end{thebibliography}